A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-18 paper/paper.Rmd
Added hgc_1979 explanation.

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (`gender`), race (`race`), age (`age_1979`), highest grade completed reported in each round of survey (`grade`), highest grade completed ever reported (`hgc`), highest grade completed in terms of years, e.g. 9th grade = 9, 3rd year college = 15, (`hgc_i`), and whether the graduate equivalency diploma is obtained (`ged`).

It is worth noting that `gender` comes from a variable called `sex` in the database. We use the term `gender` as the reports about the data on various pages in the website more commnly use `gender`. Gender, as provided in the data, is self-reported and only has two categories, "male" and "female". The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification [@kennedy2020using]. From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex [@kennedy2020using]. Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The raw data, `new_data_qnames`, contains the variables `Q1-3_A~Y_1979` and `Q1-3_A~Y_1981` which records two versions of the birth year of the respondent; this is also the case for the record of birth month (`Q1-3_A~M_1979` and `Q1-3_A~M_1981`). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

@@

The refreshed dataset has two types of highest grade completed. The first one is the highest grade completed ever reported in the database (`hgc` and `hgc_i`, for the factor and integer type, respectively). For each ID, there is only one value of `hgc` and `hgc_i`. This variable is obtained from `new_data_qnames` with name `HGC_EVER_XRND` and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be `hgc_i` and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed corresponds to value  reported in each round of the survey (`grade`), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in `new_data_qnames` as variables beginning with `Q3-4` and `HGC` with suffix of the year it was recorded. In addition the variables beginning with `HGCREV` contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years. <!--We also check the consistency between `hgc_i` and `grade`. Some problematic observations were found, for example, `grade` value is greater than `hgc_i`. We then fix this problem by replacing those `grade` that is greater than its `hgc_i`. -->

@@

Reading the section titled ["Topical Guide to the Data"](https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience) in the guide, suggests that it should be calculated based on the variable "number of weeks worked since last interview". This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (as discussed futher in Section \@ref(takeaways)).

@@

The list of variables provided in the three new datasets are as follows:

@@

  1. `id`: A unique individual's ID number.
  2. `age_1979`: The age of the individual in 1979.
  3. `gender`: Gender of the individual, FEMALE and MALE.
  4. `race` : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
  5. `hgc`: The highest grade completed.
  6. `hgc_i`: Integer value of the highest grade completed.
  7. `ged`: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.

@@

  1. `id`: A unique individual's ID number. This is the `key` of the data as we saved the data as a `tsibble` object. 
  2. `year`: The year the observation was taken. This is the `index` of the data.
  3. `wage`: The mean of the hourly wages the individual gets at each of their different jobs. The value could be a weighted or an arithmetic mean. The weighted mean is used when the information of hours of work as the weight is available. The mean hourly wage could also be a predicted value if the original value is considered influential by the robust linear regression as part of data cleaning.
  4. `age_1979`: The age of the individual in 1979.
  5. `gender`: Gender of the individual: FEMALE and MALE.
  6. `race`: Race where the individual belongs to: NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
  7. `hgc`: The highest grade completed ever.
  8. `hgc_i`: Integer value of the highest grade completed ever.
  9. `grade`: Integer value of the highest grade completed corresponding to `year`. 
  10. `ged`: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
  11. `njobs`: Number of jobs that an individual has.
  12. `hours:`: The total number of hours the individual usually works per week.
  13. `stwork`: The year when the individual starting to work.
  14. `yr_wforce`: The length of time in the workforce in years (`year` - `stwork`).
  15. `exp`: Work experience, i.e., the number of years of working. 
  16. `is_wm`: Whether the mean hourly wage is weighted mean, using the hour work as the weight, or regular/arithmetic mean. TRUE = is weighted mean. FALSE = is regular mean.
  17. `is_pred`: Whether the mean hourly wage is a predicted value of RLM or not.

@@

@@

@@

Figure \ref{fig:compare-subsets} contains summaries of corresponding variables in the two subsets. Plot (A) shows a back-back bar chart of the highest grade completed. The two sets are almost the same, but small differences remain. Plots (B) and (C) show stacked density plots of experience and log wages, respectively. The distributions are relatively close, with more differences in wages as would be expected becuase the refreshed data is not inflation-adjusted.

```{r calculate-summaries-on-subsets}
```

```{r compare-subsets, fig.cap = "Comparison of original and refreshed data: (A) highest grade completed, (B) experience and (C) log wages. Some difference in wages would be expected because the refreshed data is not inflation-adjusted, but the two sets are reasonably similar.", fig.width = 9, fig.height = 5, out.width = "100%", cache = FALSE}
```

@@

@@

>>>>

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (`gender`), race (`race`), age (`age_1979`), highest grade completed reported in each round of survey (`grade`), highest grade completed ever reported (`hgc`), highest grade completed in terms of years, e.g. 9th grade = 9, 3rd year college = 15, (`hgc_i`), highest grade completed in 1979 (`hgc_1979`) and whether the graduate equivalency diploma is obtained (`ged`).

For `gender` and `race`, we only rename the column names provided in the raw data, `new_data_qnames`. It is worth noting that `gender` comes from a variable called `sex` in the database. We use the term `gender` as the reports about the data on various pages in the website more commonly use `gender`. Gender, as provided in the data, is self-reported and only has two categories, "male" and "female". The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification [@kennedy2020using]. From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex [@kennedy2020using]. Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

`new_data_qnames`, contains the variables `Q1-3_A~Y_1979` and `Q1-3_A~Y_1981` which records two versions of the birth year of the respondent; this is also the case for the record of birth month (`Q1-3_A~M_1979` and `Q1-3_A~M_1981`). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

@@

The refreshed dataset has three types of highest grade completed. The first one is the highest grade completed ever reported in the database (`hgc` and `hgc_i`, for the factor and integer type, respectively). For each ID, there is only one value of `hgc` and `hgc_i`. This variable is obtained from `new_data_qnames` with name `HGC_EVER_XRND` and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be `hgc_i` and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed in 1979 (`hgc_1979`), the first round of the survey. We decide to include this variable in the dataset for comparison with the original data-purpose. The original data only covers the period until 1994. Hence, `hgc` in the refreshed data, which is the state of `hgc` in 2018, and the `hgc` in the original data will not be comparable.

The third one is the highest grade completed corresponds to value  reported in each round of the survey (`grade`), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in `new_data_qnames` as variables beginning with `Q3-4` and `HGC` with suffix of the year it was recorded. In addition the variables beginning with `HGCREV` contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years. <!--We also check the consistency between `hgc_i` and `grade`. Some problematic observations were found, for example, `grade` value is greater than `hgc_i`. We then fix this problem by replacing those `grade` that is greater than its `hgc_i`. -->

@@

Reading the section titled ["Topical Guide to the Data"](https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience) in the guide, suggests that it should be calculated based on the variable "number of weeks worked since last interview". This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (as discussed further in Section \@ref(takeaways)).

@@

In addition, the list of variables that we cover in those three datasets are as follows:

@@

  1. `id`: A unique individual's ID number.
  2. `age_1979`: The age of the individual in 1979.
  3. `gender`: Gender of the individual, FEMALE and MALE.
  4. `race` : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
  5. `hgc`: The highest grade completed ever.
  6. `hgc_i`: Integer value of the highest grade completed ever.
  7. `hgc_1979`: The highest grade completed in 1979 (integer value). 
  8. `ged`: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.

@@

  1. `id`: A unique individual's ID number. This is the `key` of the data as we saved the data as a `tsibble` object. 
  2. `year`: The year the observation was taken. This is the `index` of the data.
  3. `wage`: The mean of the hourly wages the individual gets at each of their different jobs. The value could be a weighted or an arithmetic mean. The weighted mean is used when the information of hours of work as the weight is available. The mean hourly wage could also be a predicted value if the original value is considered influential by the robust linear regression as part of data cleaning.
  4. `age_1979`: The age of the individual in 1979.
  5. `gender`: Gender of the individual: FEMALE and MALE.
  6. `race`: Race where the individual belongs to: NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
  7. `grade`: Integer value of the highest grade completed corresponding to `year`. 
  8. `hgc`: The highest grade completed ever.
  9. `hgc_i`: Integer value of the highest grade completed ever.
  10. `hgc_1979`: The highest grade completed in 1979 (integer value).

  11. `ged`: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
  12. `njobs`: Number of jobs that an individual has.
  13. `hours:`: The total number of hours the individual usually works per week.
  14. `stwork`: The year when the individual starting to work.
  15. `yr_wforce`: The length of time in the workforce in years (`year` - `stwork`).
  16. `exp`: Work experience, i.e., the number of years of working. 
  17. `is_wm`: Whether the mean hourly wage is weighted mean, using the hour work as the weight, or regular/arithmetic mean. TRUE = is weighted mean. FALSE = is regular mean.
  18. `is_pred`: Whether the mean hourly wage is a predicted value of RLM or not.

@@

```{r sw_wages}
```

```{r do_refreshed}
```

```{r compare-sw-do}
```

@@

```{r compare-xp-sw, fig.cap = "Density plot comparing the calculated work experience in the refreshed data, up to 1994, with that of the original data for male dropouts. Experience is shown on a square root scale to reduce the skewness. The highest concentration of observations is along the x=y axis, which means that most observations at most time points are similar.", out.width = "70%", eval=FALSE}
```

```{r plotting-sw-do, fig.cap = "Comparison of original textbook example (A) with refreshed data (B). The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which makes it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%", cache = FALSE, eval=FALSE}
```

@@

Figure \ref{fig:compare-subsets} contains summaries of corresponding variables in the two subsets.

```{r compare-subsets, fig.cap = "Comparison of original textbook example with refreshed data: (A) highest grade completed, (B) experience and (C) log wages", fig.width = 9, fig.height = 7, out.width = "100%", cache = FALSE}
```

@@

```{r compare-xp-to-yrworkforce, fig.cap = "The association between the lenght of time in the workforce and the work experience. Generally, the longer someone were in the workforce, the higher the work experience. However, some of the respondents have much lower working experience compared to their duration in the workforce. This is possible when someone ", eval=FALSE}
```

@@

```{r summaries, eval=FALSE}
```

<!-- Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data. It is also worth noting that the textbook data is reused from @MurnaneRichardJ1999DMDB’s paper. Hence, it is not directly extracted from the NLSY79 database.  @SingerJudithD2003Alda do not state the time frame covered by the data. However, from @MurnaneRichardJ1999DMDB’s paper, we find that the textbook data cover the survey period from 1979-1994. Hence, for comparison purposes, we filter the refreshed data only until 1994.

In the absence of highs chool dropouts criteria in the textbook data, we create a clearer criteria of high school dropouts as mentioned in Section \@ref(#target). Consequently, we do not have the same ID of high-school dropouts as in the textbook data. The textbook data has 888 IDs included, while the refreshed data has xx IDs. Furthermore, there are only xx IDs included both in the textbook and the refreshed data. After observing @MurnaneRichardJ1999DMDB’s paper, it is mentioned that the data is randomly drawn from the low income subsample of the NLSY. This is probably the reason why, even using some criteria, we cannot match the dropouts in the textbook and the refreshed data.

Further, the temporal variable used in the textbook data is experience, instead of year. The only explanation we found regarding this variable is that it is derived from the years of IDs’ experience since their first job. @SingerJudithD2003Alda stated that this variable reflects the specific moment in each ID labor force history associated with each wage observation. 
Since it is not clearly articulated, we decide to keep the temporal variable in the refreshed data as year. Accordingly, in this comparison, we only compare the same ID but not the same temporal variable.

Regarding the wages, the textbook data express the wages as inflation adjusted with 1990 dollars as constant [@SingerJudithD2003Alda]. However, it is not stated how exactly the wages are adjusted. Note that in the refreshed data we publish in `yowie`, we do not adjust the data with inflation because the time frame covered ranged from 1979-2018. Hence, we allow flexibility to the data user to adjust, not adjust,  or adjust it with different base years. For comparison purposes, we adjust the wages with the average Consumer Price Index For All Urban Consumers (CPI-U) we got from @fed. -->

<!-- Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data. Even though it has different temporal variables and possibly different CPI indicators used in inflation adjustment, we can still capture some of the patterns in the refreshed data.

From this comparison, we learn that we cannot make an apple to apple comparison and produce the exact same data as the textbook due to the absence of clearly articulated data extraction and derivation process. Hence, the takeaway is that it is imperative to disclose how the data derived from the initial source to make data analysis reproducible. This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.-->

### 2022-02-18 paper/paper.tex
Added hgc_1979 explanation.

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age\_1979}), highest grade completed reported in each round of survey (\texttt{grade}), highest grade completed ever reported (\texttt{hgc}), highest grade completed in terms of years, e.g.~9th grade = 9, 3rd year college = 15, (\texttt{hgc\_i}), and whether the graduate equivalency diploma is obtained (\texttt{ged}).

It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commnly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female''. The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The raw data, \texttt{new\_data\_qnames}, contains the variables \texttt{Q1-3\_A\textasciitilde{}Y\_1979} and \texttt{Q1-3\_A\textasciitilde{}Y\_1981} which records two versions of the birth year of the respondent; this is also the case for the record of birth month (\texttt{Q1-3\_A\textasciitilde{}M\_1979} and \texttt{Q1-3\_A\textasciitilde{}M\_1981}). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

The refreshed dataset has two types of highest grade completed. The first one is the highest grade completed ever reported in the database (\texttt{hgc} and \texttt{hgc\_i}, for the factor and integer type, respectively). For each ID, there is only one value of \texttt{hgc} and \texttt{hgc\_i}. This variable is obtained from \texttt{new\_data\_qnames} with name \texttt{HGC\_EVER\_XRND} and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be \texttt{hgc\_i} and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed corresponds to value reported in each round of the survey (\texttt{grade}), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in \texttt{new\_data\_qnames} as variables beginning with \texttt{Q3-4} and \texttt{HGC} with suffix of the year it was recorded. In addition the variables beginning with \texttt{HGCREV} contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years.

@@

Reading the section titled \href{https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience}{``Topical Guide to the Data''} in the guide, suggests that it should be calculated based on the variable ``number of weeks worked since last interview''. This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (as discussed futher in Section \ref{takeaways}).

@@

\caption{The stages of data cleaning from the raw data to get three datasets contained in \texttt{[CENSORED]}. ``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data. The color represents the stage of data cleaning in statistical value chain (M. P. J. van der Loo and de Jonge 2021). Pink, blue, and green represent the raw, input, and valid data, respectively.}\label{fig:flow-chart-blind}
\end{figure}

The list of variables provided in the three new datasets are as follows:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \texttt{id}: A unique individual's ID number.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual, FEMALE and MALE.
\item
  \texttt{race} : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{hgc}: The highest grade completed.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed.
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \texttt{id}: A unique individual's ID number. This is the \texttt{key} of the data as we saved the data as a \texttt{tsibble} object.
\item
  \texttt{year}: The year the observation was taken. This is the \texttt{index} of the data.
\item
  \texttt{wage}: The mean of the hourly wages the individual gets at each of their different jobs. The value could be a weighted or an arithmetic mean. The weighted mean is used when the information of hours of work as the weight is available. The mean hourly wage could also be a predicted value if the original value is considered influential by the robust linear regression as part of data cleaning.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual: FEMALE and MALE.
\item
  \texttt{race}: Race where the individual belongs to: NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{hgc}: The highest grade completed ever.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed ever.
\item
  \texttt{grade}: Integer value of the highest grade completed corresponding to \texttt{year}.
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\item
  \texttt{njobs}: Number of jobs that an individual has.
\item
  \texttt{hours:}: The total number of hours the individual usually works per week.
\item
  \texttt{stwork}: The year when the individual starting to work.
\item
  \texttt{yr\_wforce}: The length of time in the workforce in years (\texttt{year} - \texttt{stwork}).
\item
  \texttt{exp}: Work experience, i.e., the number of years of working.
\item
  \texttt{is\_wm}: Whether the mean hourly wage is weighted mean, using the hour work as the weight, or regular/arithmetic mean. TRUE = is weighted mean. FALSE = is regular mean.
\item
  \texttt{is\_pred}: Whether the mean hourly wage is a predicted value of RLM or not.
\end{enumerate}

@@

Figure \ref{fig:compare-subsets} contains summaries of corresponding variables in the two subsets. Plot (A) shows a back-back bar chart of the highest grade completed. The two sets are almost the same, but small differences remain. Plots (B) and (C) show stacked density plots of experience and log wages, respectively. The distributions are relatively close, with more differences in wages as would be expected becuase the refreshed data is not inflation-adjusted.

@@

\caption{Comparison of original and refreshed data: (A) highest grade completed, (B) experience and (C) log wages. Some difference in wages would be expected because the refreshed data is not inflation-adjusted, but the two sets are reasonably similar.}\label{fig:compare-subsets}
\end{figure}

>>>>

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age\_1979}), highest grade completed reported in each round of survey (\texttt{grade}), highest grade completed ever reported (\texttt{hgc}), highest grade completed in terms of years, e.g.~9th grade = 9, 3rd year college = 15, (\texttt{hgc\_i}), highest grade completed in 1979 (\texttt{hgc\_1979}) and whether the graduate equivalency diploma is obtained (\texttt{ged}).

For \texttt{gender} and \texttt{race}, we only rename the column names provided in the raw data, \texttt{new\_data\_qnames}. It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commonly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female''. The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

\texttt{new\_data\_qnames}, contains the variables \texttt{Q1-3\_A\textasciitilde{}Y\_1979} and \texttt{Q1-3\_A\textasciitilde{}Y\_1981} which records two versions of the birth year of the respondent; this is also the case for the record of birth month (\texttt{Q1-3\_A\textasciitilde{}M\_1979} and \texttt{Q1-3\_A\textasciitilde{}M\_1981}). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

The refreshed dataset has three types of highest grade completed. The first one is the highest grade completed ever reported in the database (\texttt{hgc} and \texttt{hgc\_i}, for the factor and integer type, respectively). For each ID, there is only one value of \texttt{hgc} and \texttt{hgc\_i}. This variable is obtained from \texttt{new\_data\_qnames} with name \texttt{HGC\_EVER\_XRND} and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be \texttt{hgc\_i} and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed in 1979 (\texttt{hgc\_1979}), the first round of the survey. We decide to include this variable in the dataset for comparison with the original data-purpose. The original data only covers the period until 1994. Hence, \texttt{hgc} in the refreshed data, which is the state of \texttt{hgc} in 2018, and the \texttt{hgc} in the original data will not be comparable.

The third one is the highest grade completed corresponds to value reported in each round of the survey (\texttt{grade}), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in \texttt{new\_data\_qnames} as variables beginning with \texttt{Q3-4} and \texttt{HGC} with suffix of the year it was recorded. In addition the variables beginning with \texttt{HGCREV} contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years.

@@

Reading the section titled \href{https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience}{``Topical Guide to the Data''} in the guide, suggests that it should be calculated based on the variable ``number of weeks worked since last interview''. This would remove periods of unemployment which makes sense when measuring experience while actually working. Since it is only measured since the last interview, this needs to be cumulated for each survey year. This produced results more similar to the original data (as discussed further in Section \ref{takeaways}).

@@

\caption{The stages of data cleaning from the raw data to get three datasets contained in \texttt{[CENSORED]}. ``\# of individuals'' means the number of respondents included in each stage, while ``\# of observations'' means the number of rows in the data. The color represents the stage of data cleaning in statistical value chain (van der Loo and de Jonge 2021). Pink, blue, and green represent the raw, input, and valid data, respectively.}\label{fig:flow-chart-blind}
\end{figure}

In addition, the list of variables that we cover in those three datasets are as follows:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \texttt{id}: A unique individual's ID number.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual, FEMALE and MALE.
\item
  \texttt{race} : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{hgc}: The highest grade completed ever.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed ever.
\item
  \texttt{hgc\_1979}: The highest grade completed in 1979 (integer value).
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \texttt{id}: A unique individual's ID number. This is the \texttt{key} of the data as we saved the data as a \texttt{tsibble} object.
\item
  \texttt{year}: The year the observation was taken. This is the \texttt{index} of the data.
\item
  \texttt{wage}: The mean of the hourly wages the individual gets at each of their different jobs. The value could be a weighted or an arithmetic mean. The weighted mean is used when the information of hours of work as the weight is available. The mean hourly wage could also be a predicted value if the original value is considered influential by the robust linear regression as part of data cleaning.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual: FEMALE and MALE.
\item
  \texttt{race}: Race where the individual belongs to: NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{grade}: Integer value of the highest grade completed corresponding to \texttt{year}.
\item
  \texttt{hgc}: The highest grade completed ever.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed ever.
\item
  \texttt{hgc\_1979}: The highest grade completed in 1979 (integer value).
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\item
  \texttt{njobs}: Number of jobs that an individual has.
\item
  \texttt{hours:}: The total number of hours the individual usually works per week.
\item
  \texttt{stwork}: The year when the individual starting to work.
\item
  \texttt{yr\_wforce}: The length of time in the workforce in years (\texttt{year} - \texttt{stwork}).
\item
  \texttt{exp}: Work experience, i.e., the number of years of working.
\item
  \texttt{is\_wm}: Whether the mean hourly wage is weighted mean, using the hour work as the weight, or regular/arithmetic mean. TRUE = is weighted mean. FALSE = is regular mean.
\item
  \texttt{is\_pred}: Whether the mean hourly wage is a predicted value of RLM or not.
\end{enumerate}

@@

Figure \ref{fig:compare-subsets} contains summaries of corresponding variables in the two subsets.

@@

\caption{Comparison of original textbook example with refreshed data: (A) highest grade completed, (B) experience and (C) log wages}\label{fig:compare-subsets}
\end{figure}
